\Title

### 2021-05-27 latex/paper/main.tex
Finished editing the words that Manfredi wanted me to edit

We can extend this idea to find the nominal pendulum associated with Wang's
acrobot, whose physical parameters are provided in Table
\ref{tab:acrobot-parameters}.
Setting \(I = 0\) in \eqref{eqn:acrobot-constraint} for this acrobot
yields the following mechanical energy of the nominal pendulum:
\[
    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
    .
\]
Despite the change in acrobot model, the level set \(E_\pi\) with energy 
\(E(\pi,0)\) remains the boundary between oscillations and rotations
for the nominal pendulum.

@@

Setting \(\bar{q}_a = 1\), we experimentally determined that \(|I| = 10\) is a
viable control parameter, so this is the value we will use for all simulations
and experiments.
In other words, our injection VNHC is \eqref{eqn:acrobot-constraint} with 
\(I = 10\) while our dissipation VNHC is \eqref{eqn:acrobot-constraint} with 
\(I = -10\).

\begin{table}
    \centering
    \caption{Physical parameters for the real acrobot.}
    \label{tab:acrobot-parameters}
    \resizebox{\columnwidth}{!}{%
    \begin{tabular}{ccccccccc}
        \toprule
        $m_u$ & $m_a$ & $l_u$ & $l_a$ & $l_{c_u}$ & $l_{c_a}$ & $J_u$ & $J_a$ & $g$ \\
        \midrule
        0.2112 & 0.1979 & % mass
        0.148 & 0.145 & % total length
        0.073 & 0.083 & % COM distance
        0.00129 & 0.00075 & % moment of inertia
        9.81 \\ % gravity
        \bottomrule
    \end{tabular}
    } % resizebox
\end{table}

@@

We simulated the acrobot constrained by our injection VNHC
with initial condition \((q_u,p_u) = \left(\pi/32,0 \right)\) for \(30\)
seconds, and plotted the resulting orbit in Figure \ref{fig:acrobot-in-orbit}.

The level set \(E_\pi\) with energy \(E(\pi,0)\) is outlined in black.
Recall that this marks the boundary between oscillations and rotations of the
nominal pendulum.
The points where the orbit exits \(E_\pi\) are marked with black stars,
with the final departure marked by a red star.
Interestingly, our choice of \(I\) is large enough that we observe significant
differences between the nominal pendulum and the constrained dynamics:
\(E_\pi\) intersects the \(p_u\)-axis at \(\abs{p_u} \approx 0.18\), yet the
constrained acrobot begins rotating once it hits the
\(p_u\)-axis at \(\abs{p_u} \approx 0.16\). 
This indicates that higher values of \(I\) enable the acrobot to gain energy
faster and begin rotating sooner, so long as the actuator does not saturate.

@@

We initialized our acrobot with a rotation at \((q_u,p_u) = (0,0.18)\) and
simulated the constrained dynamics (under the dissipation VNHC) for \(30\)
seconds.
The resulting orbit is plotted in Figure \ref{fig:acrobot-diss-orbit}. 
As expected, the acrobot slows down over time.
We highlight the locations where the orbit crossed the set \(E_\pi\) by
black stars, with the final crossing in red.
After this final crossing, the acrobot ceased rotating and its oscillations
decayed to zero.

@@

Recall from \ref{sec:energy-reg} that one can use a supervisor to stabilize
oscillations by appropriately toggling between injection and dissipation VNHCs.
Figure \ref{fig:acrobot-osc-reg} shows  stabilizing an
oscillation with body angle \(q_\text{des} = \pi/2\) and a 
\(5\%\) hysteresis, meaning \(\delta = 0.05\).
The supervisor reevaluated its choice of VNHC at each black star;  
the red contour corresponds to the part of the orbit where the supervisor kept
the leg extended, because the oscillation was within tolerance of
\(q_\text{des}\).
The solid black line is the desired oscillation, and the dashed black
lines show the hysteresis around that orbit.

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_osc_orbit_1.png}
        \caption{Stabilizing an oscillation from below.}
        \label{fig:acrobot-osc-reg-in}
    \end{subfigure}
    \begin{subfigure}[t]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_osc_orbit_2.png}
        \caption{Stabilizing an oscillation from above.}
        \label{fig:acrobot-osc-reg-diss}
    \end{subfigure}
    \caption{Using a supervisor to stabilize the oscillation with peak angle
    \(q_\text{des} = \pi/2\).}
    \label{fig:acrobot-osc-reg}
\end{figure}

\subsection{Rotation Regulation}
One can also use a supervisor to stabilize rotations through the mechanism
described in Section \ref{sec:energy-reg}.
Rotation regulation for the acrobot is demonstrated in 
Figure \ref{fig:acrobot-rot-reg}, where the supervisor stabilizes 
\(p_\text{des} = 0.19\) with a \(2\%\) hysteresis \(\delta = 0.02\).
The supervisor changed its choice of VNHC at each black star until it was within
range of \(p_\text{des}\), at which point it extended the leg (shown in red).

In Figure \ref{fig:acrobot-rot-reg-in} the acrobot was initialized at 
\((q_u,p_u) = (\pi/2,0)\); here the supervisor injected energy until
the orbit hit the \(p_u\)-axis near \(p_\text{des}\).
In Figure \ref{fig:acrobot-rot-reg-diss} the acrobot was initialized at the
faster rotation \((q_u,p_u) = (0, 0.23)\); here the supervisor dissipated energy.
In both cases, the desired rotation was stabilized correctly.

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_rot_orbit_1.png}
        \caption{Stabilizing a rotation from below.}
        \label{fig:acrobot-rot-reg-in}
    \end{subfigure}
    \begin{subfigure}[t]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_rot_orbit_2.png}
        \caption{Stabilizing a rotation from above.}
        \label{fig:acrobot-rot-reg-diss}
    \end{subfigure}
    \caption{Using a supervisor to stabilize the rotation with maximal momentum
    \(p_\text{des} = 0.23\).}
    \label{fig:acrobot-rot-reg}
\end{figure}

@@

\begin{enumerate}
    \item \textbf{Baseline Test:} 
    we initialized the acrobot at 
    \((q_u,p_u) \approx \left(\pi/8,0\right)\). 
    The resulting orbit in Figure \ref{fig:acrobot-unperturbed-orbit} shows
    that SUGAR clearly gaining energy over time.
    Its motion looks similar to that of the energy injection simulation (Figure
    \ref{fig:acrobot-in-orbit}), though its energy gain ceases once it reaches a
    rotation with energy \(E(0,0.195)\).

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{acrobot_unperturbed_orbit.png}
    \caption{SUGAR's orbit during the baseline energy injection test.}
    \label{fig:acrobot-unperturbed-orbit}
\end{figure}

\item \textbf{Stop Test:}
    we initialized SUGAR at 
    \((q_u,p_u) \approx \left(\pi,0\right)\), let it run for 15
    seconds, then stopped it at the bottom of its arc.
    The resulting orbit is shown in Figure \ref{fig:acrobot-stopped-orbit}.
    The blue rotation curve corresponds to the orbit before the disturbance,
    while the red spiral confirms that the acrobot begins oscillating after it
    is stopped.  
    Despite the disturbance, it gains energy and eventually starts rotating
    again.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{acrobot_stopped_orbit.png}
    \caption{SUGAR's orbit before (blue) and after (red) stopping.}
    \label{fig:acrobot-stopped-orbit}
\end{figure}

\item \textbf{Push Test:}
    to see how SUGAR responds when pushed, we allowed
    it to rotate undisturbed for 15 seconds and then pushed it in its
    direction of motion.
    The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that the acrobot,
    when pushed, rotates with energy \(E(0,-0.22)\), but then slows down until
    it reaches a rotation with energy \(E(0,-0.195)\).
    We repeated this test by pushing SUGAR against its direction of motion.
    The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that the acrobot
    readily changes direction, and quickly achieves its maximum speed
    with energy \(E(0,0.195)\).
\end{enumerate}

In simulation, the acrobot was able to gain energy even when initialized
with energy \(E(q_u,p_u) > E(0,0.195)\). 
The baseline and push tests suggest that our VNHC injects energy into SUGAR
only on \(\mathcal{O}_2(0.195)\).
This difference between simulation and implementation is likely due to
friction.

\begin{figure}
    \centering
    \begin{subfigure}[ht]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_fpush_orbit.png}
        \caption{The forwards push test.}
        \label{fig:acrobot-fpush-orbit}
    \end{subfigure}
    \begin{subfigure}[ht]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_rpush_orbit.png}
        \caption{The reverse push test.}
        \label{fig:acrobot-rpush-orbit}
    \end{subfigure}
    \caption{SUGAR's orbit before (blue) and after (red) pushing.}
\end{figure}

@@

In this article we redeveloped the framework of virtual nonholonomic
constraints for underactuated mechanical systems with arbitrary degree of
underactuation.
We applied this framework towards the acrobot, and designed a constraint which
emulates giant motion from gymnastics.
We proved this constraint will inject energy into the simplified acrobot whose
limbs are massless rods of equal length, with equal masses at the tips.
We then performed simulations and physical experiments of this VNHC.
These demonstrated that virtual nonholonomic constraints are capable of
injecting and dissipating energy in a robust manner, all while producing
realistic biological motion.

>>>>

We can extend this idea to find the nominal pendulum associated with Wang's
acrobot, whose physical parameters are provided in Table
\ref{tab:acrobot-parameters}.
Setting \(I = 0\) in \eqref{eqn:acrobot-constraint} for this acrobot
yields the following mechanical energy of the nominal pendulum:
\[
    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
    .
\]
Despite the change in model, the level set \(E_\pi\) with energy 
\(E(\pi,0)\) remains the boundary between oscillations and rotations
for this nominal pendulum.

@@

Setting \(\bar{q}_a = 1\), we experimentally determined that \(|I| = 10\) is a
viable control parameter, so this is the value we will use for all simulations
and experiments.
In other words, our \textit{injection VNHC} is \eqref{eqn:acrobot-constraint} with 
\(I = 10\) while our \textit{dissipation VNHC} is \eqref{eqn:acrobot-constraint} with 
\(I = -10\).

\begin{table}
    \centering
    \caption{Physical parameters for the real acrobot.}
    \label{tab:acrobot-parameters}
    \resizebox{\columnwidth}{!}{%
    \begin{tabular}{ccccccccc}
        \toprule
        $m_u$ & $m_a$ & $l_u$ & $l_a$ & $l_{c_u}$ & $l_{c_a}$ & $J_u$ & $J_a$ & $g$ \\
        (kg) & (kg) & (m) & (m) & (m) & (m) & (kg\(\cdot\)m\(^2\)) &
        (kg\(\cdot\)m\(^2\)) & (m/s\(^2\)) \\
        \midrule
        0.2112 & 0.1979 & % mass
        0.148 & 0.145 & % total length
        0.073 & 0.083 & % COM distance
        0.00129 & 0.00075 & % moment of inertia
        9.81 \\ % gravity
        \bottomrule
    \end{tabular}
    } % resizebox
\end{table}

@@

In simulation, we stabilized the injection VNHC for the acrobot using the
controller \eqref{eqn:stabilizing-controller}.
We initialized the acrobot on the constraint manifold
with initial condition \((q_u,p_u) = \left(\pi/32,0 \right)\) and simulated the
constrained system for \(30\) seconds.
The resulting orbit (on the constraint manifold) is plotted in Figure
\ref{fig:acrobot-in-orbit}.

The level set \(E_\pi\) with energy \(E(\pi,0)\) is outlined in black.
Recall that this level set is the boundary between oscillations and rotations of the
nominal pendulum.
The points where the orbit exits \(E_\pi\) are marked with black stars,
with the final departure marked by a red star.
Interestingly, our choice of \(I\) is large enough that we observe significant
differences between the nominal pendulum and the constrained dynamics:
\(E_\pi\) intersects the \(p_u\)-axis at \(\abs{p_u} \approx 0.18\), yet the
constrained acrobot begins rotating once it hits the
\(p_u\)-axis at \(\abs{p_u} \approx 0.16\). 
This indicates that higher values of \(I\) enable the acrobot to gain energy
faster and begin rotating sooner, so long as the actuator does not saturate.

@@

In simulation, we stabilized the dissipation VNHC and initialized the acrobot on
the constraint manifold with a rotation \((q_u,p_u) = (0,0.18)\).
We simulated the constrained system for \(30\) seconds and plotted the
resulting orbit in Figure \ref{fig:acrobot-diss-orbit}. 
As expected, the acrobot slows down over time.
We highlight the locations where the orbit crossed the set \(E_\pi\) by
black stars, with the final crossing in red.
After this final crossing, the acrobot ceased rotating and its oscillations
decayed to zero.

@@

Recall from \ref{sec:energy-reg} that one can use a supervisor to stabilize
oscillations by appropriately toggling between injection and dissipation VNHCs
whenever the orbit of the acrobot crosses the \(q_u\)-axis.
Figure \ref{fig:acrobot-osc-reg} shows the supervisor stabilizing an
oscillation with body angle \(q_\text{des} = \pi/2\) and a 
\(5\%\) hysteresis, meaning \(\delta = 0.05\).
The supervisor reevaluated its choice of VNHC at each black star;  
the red contour corresponds to the part of the orbit where the supervisor kept
the leg extended, because the oscillation was within tolerance of
\(q_\text{des}\).
The solid black line is the desired oscillation, and the dashed black
lines show the hysteresis around that orbit.

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_osc_orbit_1.png}
        \caption{Stabilizing an oscillation from below.}
        \label{fig:acrobot-osc-reg-in}
    \end{subfigure}
    \begin{subfigure}[t]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_osc_orbit_2.png}
        \caption{Stabilizing an oscillation from above.}
        \label{fig:acrobot-osc-reg-diss}
    \end{subfigure}
    \caption{Using a supervisor to stabilize the oscillation with peak angle
    \(q_\text{des} = \pi/2\). The desired oscillation is depicted with a solid
    black line.}
    \label{fig:acrobot-osc-reg}
\end{figure}

\subsection{Rotation Regulation}
One can also use a supervisor to stabilize rotations through the mechanism
described in Section \ref{sec:energy-reg}, where the supervisor toggles between
injection and dissipation VNHCs at each crossing of the \(p_u\)-axis.
Rotation regulation for the acrobot is demonstrated in 
Figure \ref{fig:acrobot-rot-reg}, where the supervisor stabilizes 
\(p_\text{des} = 0.19\) with a \(2\%\) hysteresis \(\delta = 0.02\).
The supervisor evaluated its choice of VNHC at each black star.
Once it was within range of \(p_\text{des}\) it extended the legs completely,
the orbit of which is shown in red.

In Figure \ref{fig:acrobot-rot-reg-in} the acrobot was initialized at 
the oscillation \((q_u,p_u) = (\pi/2,0)\); here the supervisor injected energy
until the orbit hit the \(p_u\)-axis near \(p_\text{des}\).
In Figure \ref{fig:acrobot-rot-reg-diss} the acrobot was initialized at the
(fast) rotation \((q_u,p_u) = (0, 0.23)\); here the supervisor dissipated energy.
In both cases, the desired rotation was stabilized correctly.

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_rot_orbit_1.png}
        \caption{Stabilizing a rotation from below.}
        \label{fig:acrobot-rot-reg-in}
    \end{subfigure}
    \begin{subfigure}[t]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_rot_orbit_2.png}
        \caption{Stabilizing a rotation from above.}
        \label{fig:acrobot-rot-reg-diss}
    \end{subfigure}
    \caption{Using a supervisor to stabilize the rotation with maximal momentum
    \(p_\text{des} = 0.23\). The desired rotation is depicted with a solid black
    line.}
    \label{fig:acrobot-rot-reg}
\end{figure}

@@

\begin{enumerate}
    \item \textbf{Baseline Test:} 
    we initialized SUGAR at 
    \((q_u,p_u) \approx \left(\pi/8,0\right)\). 
    The resulting orbit in Figure \ref{fig:acrobot-unperturbed-orbit} shows
    that SUGAR clearly gaining energy over time.
    Its motion looks similar to that of the energy injection simulation (Figure
    \ref{fig:acrobot-in-orbit}), though its energy gain ceases once it reaches a
    rotation with energy \(E(0,0.195)\).
    The energy gain likely ceases because of friction at the pivot, which was
    not modelled in simulation.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{acrobot_unperturbed_orbit.png}
    \caption{Baseline Test: SUGAR's baseline energy injection orbit.}
    \label{fig:acrobot-unperturbed-orbit}
\end{figure}

\item \textbf{Perturbation Test 1:}
    we initialized SUGAR at 
    \((q_u,p_u) \approx \left(\pi,0\right)\), let it run for 15
    seconds, then introduced a rod as it passed through the bottom of
    its arc.
    This caused a collision which stopped SUGAR in its tracks, at which
    point we immediately removed the rod so SUGAR could continue
    unperturbed.
    The resulting orbit is shown in Figure \ref{fig:acrobot-stopped-orbit}.
    The blue rotation curve corresponds to the orbit before the disturbance,
    while the red spiral confirms that SUGAR begins oscillating after it
    was stopped.  
    After the collision, SUGAR gains energy and eventually starts
    rotating again.

\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{acrobot_stopped_orbit.png}
    \caption{Perturbation Test 1: SUGAR's orbit before (blue) and after (red) stopping.}
    \label{fig:acrobot-stopped-orbit}
\end{figure}

\item \textbf{Perturbation Test 2:}
    to see how SUGAR responds when pushed, we allowed
    it to rotate unperturbed for 15 seconds and then pushed it in its
    direction of motion.
    The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that SUGAR,
    when pushed, rotates with energy \(E(0,-0.22)\), but then slows down until
    it reaches a rotation with energy \(E(0,-0.195)\).
    We repeated this test by pushing SUGAR against its direction of motion.
    The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that it
    readily changes direction, and quickly achieves its maximum speed
    with energy \(E(0,0.195)\).
\end{enumerate}

In simulation, the acrobot was able to gain energy even when initialized
with energy \(E(q_u,p_u) > E(0,0.195)\). 
The baseline and push tests suggest that our VNHC injects energy into SUGAR
only on \(\mathcal{O}_2(0.195)\).
This difference between simulation and implementation is likely due to
friction, as well as timing errors incurred by the PID controller in the servo
motor.

\begin{figure}
    \centering
    \begin{subfigure}[ht]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_fpush_orbit.png}
        \caption{The forwards push test.}
        \label{fig:acrobot-fpush-orbit}
    \end{subfigure}
    \begin{subfigure}[ht]{0.49\linewidth}
        \includegraphics[width=\linewidth]{acrobot_rpush_orbit.png}
        \caption{The reverse push test.}
        \label{fig:acrobot-rpush-orbit}
    \end{subfigure}
    \caption{Perturbation Test 2: SUGAR's orbit before (blue) and after (red) pushing.}
\end{figure}

@@

In this article we apply the framework of virtual nonholonomic constraints
towards the acrobot, and designed a constraint which emulates giant motion from
gymnastics.
We proved this constraint will inject energy into the simplified acrobot whose
limbs are massless rods of equal length, with equal masses at the tips.
We then performed simulations and physical experiments on a real acrobot robot.
These demonstrated that virtual nonholonomic constraints are capable of
injecting and dissipating energy in a robust manner, all while producing
realistic biologically-inspired motion.
